Environmental and geographic variables are effective surrogates for genetic variation in conservation planning

### 2017-10-18 code/rmarkdown/si-preamble.tex
+ article within page pnas limits
+ assorted changes
+ fix supp info figures + tables + reorder si figures and tables + update si figures and table numbers in main tex
+ fix bug causing refs not to appear in article
minor tweaks to ms changes to analysis that were not pushed clean up documentation

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec,natbib,pdfpages,lineno,booktabs}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage[doublespacing]{setspace}

@@

% insert S before figure and table caption
\pagenumbering{gobble}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

>>>>

% load packages
\usepackage{amsmath,amsfonts,makecell,tocloft,titling,natbib,hyperref,booktabs,caption}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}

@@

% insert S before figure and table caption
\pagenumbering{gobble}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\captionsetup[figure]{labelfont={bf},labelformat={default},labelsep=period,name={Fig.}}
\captionsetup[table]{labelfont={bf},labelformat={default},labelsep=period,name={Table}}

### 2017-10-18 code/rmarkdown/supporting_information.Rmd
minor tweaks to ms changes to analysis that were not pushed clean up documentation

---
fontsize: 11pt
documentclass: article
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble.tex, si-preamble.tex]
---
# Supporting Information
All data and code is stored in an online repository (\url{www.github.com/jeffreyhanson/genetic-surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=(2.5*4), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environmental variation it also secures. Points represent values associated with randomly generated prioritizations. Blue lines indicate average trends computed using a generalised linear model with a logit link. Each panel shows data for a different species.'}
# format data
curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
curr.DF2$surrogate.held <- 0
curr.DF2$genetic.held <- 0.9

# make plot
suppressWarnings(
  ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
    theme_classic() +
    theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
      panel.margin.x=unit(1, 'lines')) +
    xlab('Environmental variation secured (%)') +
    ylab('Adaptive genetic variation secured (%)') +
    geom_hex() +
    suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
      se=FALSE, fill='blue', alpha=0.8)) +
    geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
    facet_wrap(~ Species, ncol=4)
)
```

```{r, include=FALSE}
calc.fig.height3 <- function(x) {
  return((((x-1) %/% 4)+1)*1.5)
}

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritization and the proportion of geographic variation it also secures. Conventions are detailed in Figure S28.'}
# format data
curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
curr.DF2$surrogate.held <- 0
curr.DF2$genetic.held <- 0.9

# make plot
suppressWarnings(
  ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
    theme_classic() +
    theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
      panel.margin.x=unit(1, 'lines')) +
    xlab('Geographic variation secured (%)') +
    ylab('Neutral genetic variation secured (%)') +
    geom_hex() +
    suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
      se=FALSE, fill='blue', alpha=0.8)) +
    geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
    facet_wrap(~ Species, ncol=4)
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(6.5*0.8), out.width='3.42in', height='3.9in', fig.cap='Planning unit selection frequencies. Squares represent planning units, and their colors indicate their selection frequency. Each column of panels shows solutions generated using different targets. Each row of panels shows solutions generated under different scenarios. Note that the selection frequencies for the single-species prioritizations are based on their frequency among prioritizations generated for each species. The selection frequencies for the multi-species prioritizations are each based on a single optimal prioritization for all species and so are binary.'}
# prepare data for plotting
grid.sub.FPLY <- grid.sub.PLY
for (i in seq_along(single.spp.prioritisations[[1]]))
  grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
  grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(multi.spp.prioritisations))
  grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
for (i in seq_along(multi.spp.prioritisations.with.cost))
  grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
grid.sub.FPLY <- spFortify(grid.sub.FPLY)
grid.sub.FPLY <- grid.sub.FPLY %>% gather(
    Combination,
    `Selection.Frequency`,
    `Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
  ) %>% mutate(
    Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
    Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
  ) %>% mutate(
    Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
    Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
    Context=factor(
      gsub(' (', '\n(', Context, fixed=TRUE),
      levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
        "Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
  )

@@

```{r, echo=FALSE}
write.table(loci_classification.DF, 'article/Table_S1.csv', sep=',', row.names=FALSE)
```

>>>>

---
fontsize: 11pt
documentclass: article
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble.tex, si-preamble.tex]
---
# Supporting Information
To replicate this study, all data and code are stored in an online repository (\url{www.github.com/jeffreyhanson/genetic-surrogates}) and are available under the digital object identifier (DOI) \texttt{10.5281/zenodo.843625}. See the \textit{README} file for instructions.

@@

```{r, include=FALSE}
calc.fig.height31 <- function(x) {
  return((((x-1) %/% 4)+1)*2.0)
}
```

```{r, include=FALSE}
calc.fig.height32 <- function(x) {
  return((((x-1) %/% 4)+1)*1.75)
}
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=11, fig.height=calc.fig.height31(n_distinct(env.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environmental variation it also secures. Points represent values associated with randomly generated prioritizations. Lines show trends fit using generalised linear models with a logit link functions. Each panel corresponds to a different species.'}
# format data
curr.DF <- env.surrogacy.DF %>% 
  filter(!is.na(genetic.held)) %>% 
  mutate(Species=gsub('_', '~', Species)) %>%
  left_join(env.surrogate.spp.DF %>% 
            filter(!is.na(r2CU)) %>% 
            mutate(Species=gsub('_', '~', Species)), by='Species')
curr.DF$R2_repr <- round(curr.DF$r2CU,2)
curr.DF$R2_repr <- sapply(curr.DF$R2_repr, function(x) {
  as.character(as.expression(substitute(italic(R)^2~"="~r2, list(r2=x))))})
curr.DF$Species_repr <- paste0('atop(italic(', curr.DF$Species, '),(', curr.DF$R2_repr, '))')

# make plot
suppressWarnings(
  ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
    theme_classic() +
    theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
      panel.margin.x=unit(1, 'lines'), legend.position='bottom') +
    xlab('Environmental variation secured (%)') +
    ylab('Adaptive genetic variation secured (%)') +
    stat_binhex(aes(fill = log(..count..))) +
    scale_fill_continuous(name='log(count)') +
    suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
      se=FALSE, color='#ffb100')) +
    facet_wrap(~ Species_repr, ncol=4, labeller = label_parsed)
)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=11, fig.height=calc.fig.height32(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritization and the proportion of geographic variation it also secures. Conventions are detailed in Figure S28.'}
# format data
curr.DF <- geo.surrogacy.DF %>% 
  filter(!is.na(genetic.held)) %>% 
  mutate(Species=gsub('_', '~', Species)) %>%
  left_join(geo.surrogate.spp.DF %>% 
            filter(!is.na(r2CU)) %>% 
            mutate(Species=gsub('_', '~', Species)), by='Species')
curr.DF$R2_repr <- round(curr.DF$r2CU,2)
curr.DF$R2_repr <- sapply(curr.DF$R2_repr, function(x) {
  as.character(as.expression(substitute(italic(R)^2~"="~r2, list(r2=x))))})
curr.DF$Species_repr <- paste0('atop(italic(', curr.DF$Species, '),(', curr.DF$R2_repr, '))')

# make plot
suppressWarnings(
  ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
    theme_classic() +
    theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
      panel.margin.x=unit(1, 'lines'), legend.position='bottom') +
    xlab('Geographic variation secured (%)') +
    ylab('Neutral genetic variation secured (%)') +
    stat_binhex(aes(fill = log(..count..))) +
    scale_fill_continuous(name='log(count)') +
    suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
      se=FALSE, color='#ffb100')) +
    facet_wrap(~ Species_repr, ncol=4, labeller = label_parsed)
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(6.5*0.8), out.width='3.42in', height='3.9in', fig.cap='Planning unit selection frequencies. Squares represent planning units, and their colors indicate their selection frequency. Each column of panels shows solutions generated using different targets. Each row of panels shows solutions generated under different scenarios. Note that the selection frequencies for the single-species prioritizations are based on their frequency among prioritizations generated for each species. The selection frequencies for the multi-species prioritizations are each based on a single optimal prioritization for all species and so are binary.'}
# prepare data for plotting
grid.sub.FPLY <- grid.sub.PLY
for (i in seq_along(single.spp.prioritisations[[1]]))
  grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
  grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(multi.spp.prioritisations))
  grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
for (i in seq_along(multi.spp.prioritisations.with.cost))
  grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
grid.sub.FPLY <- spFortify(grid.sub.FPLY)
grid.sub.FPLY <- grid.sub.FPLY %>% gather(
    Combination,
    `Selection.Frequency`,
    `Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
  ) %>% mutate(
    Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
    Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
  ) %>% mutate(
    Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
    Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
    Context=factor(
      gsub(' (', '\n(', Context, fixed=TRUE),
      levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
        "Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
  ) %>% mutate(Context = gsub('opportunity', 'acquisition', Context, fixed=TRUE))

@@

```{r, echo=FALSE}
loci_classification.DF %>%
  mutate(Species = gsub('\\textit{', '', Species, fixed=TRUE)) %>%
  mutate(Species = gsub('}', '', Species, fixed=TRUE)) %>%
  write.table('article/Table_S1.csv', sep=',', row.names=FALSE)
```
